Precise Asymptotics of Transferring from Different Regression in High Dimensions

### 2021-06-12 revision_journal/multitask.tex
restructure

%\begin{enumerate}
%	\item $X = Z \Sigma^{1/2} \in \real^{n\times p}$ for a positive semidefinite matrix $\Sigma\in \R^{p\times p}$, and every entry of $Z \in \real^{n \times p}$ is drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$.
%	\item every entry of $\varepsilon^{(i)} \in \real^{n \times t}$ is drawn indepdently from a one dimensional distribution with zero mean, variance $\sigma^2$, and bounded moment up to any order.\footnote{There exists a fixed function $C: \mathbb{N} \rightarrow \real^+$ such that for any $k \in \N$, the $k$-th moment is bounded by $C(k)$.}
%\end{enumerate}

@@

\paragraph{How does hard parameter sharing scale with sample size $n$?}
Obviously, the concentration error decreases with $n$.
First, we consider the variance of $\hat{\beta}_i^{\MTL}$, which is $\sigma^2\norm{a_i^{\star}} \bigtr{\Sigma (X^{\top} X)^{-1}}$?
It turns out that this quantity converges to a fixed limit in the high-dimensional setting, which is formally stated in the following assumption.

@@

\paragraph{Sample efficiency}
First, we validate the result of Example \ref{ex_same_cov}.
Figure \ref{fig_same_cov} shows the average prediction loss over ten tasks as we increase the number of samples per-task from $400$ to $2000$.
In all the parameter settings, our results estimate the empirical losses accurately.
We also observe a trend that the average prediction loss increases as we increase distance $d$ from $0.1$ to $0.2$.
Our work explains the differences between these two settings since $d^2 = 0.1^2$ is always smaller than $\frac{\sigma^2 p}{n - p}$, but $d^2 = 0.2^2$ is not.
%Second, when $d = 0.1$, we have that $d^2 \le \frac{\sigma^2 p}{n - p}$ for all values of $n$, hence the average prediction loss of hard parameter sharing is always lower than STL.
Indeed, we observe a crossover point between hard parameter sharing and STL.
Finally, for $d = 0.2$, looking horizontally, we find that HPS requires fewer samples per-task than STL to achieve the same loss level. %\FY{I do not quite understand this sentence about "3x fewer", because how much data needed depends on $d$ and the prediction loss level we are looking at. For example, at the cross point, this ratio is 1. }

\paragraph{Sample size ratio}
Second, we validate the result of Example \ref{ex_sample_ratio}.
Figure \ref{fig_size} shows task two's prediction loss  as we increase the sample ratio $n_1 / n_2$ from $1/10$ to $7/10$.
%Again, our estimates are accurate compared to the empirical losses.
We consider a regime where task two consists of $80,000$ samples, and task one's sample size varies from $8,000$ to $56,000$.
The task-specific variance (which scales with model distance) is $d = 0.2$, the noise variance is $\sigma^2 = 4^2$, and the shared signal variance is $1$. We observe that as we increase the sample ratio, task two's prediction loss decreases initially but later will increase when the sample ratio is above a certain level.
On the other hand, when $d = 0.15$, task two's prediction loss decreases faster.
Intuitively, this is because bias increases less for smaller $d^2$.
%Our result from Corollary \ref{cor_MTL_loss} explains this trend.

\paragraph{Covariate shift}
Finally, we validate the result of Example \ref{ex_covshift}.
Figure \ref{fig_covariate} shows task two's prediction loss as we increase task one's sample size.
Recall that $\lambda$ measures the severity of covariate shifts---a larger $\lambda$ means a larger covariate shift.
We indeed observe the dichotomy in Example \ref{ex_covshift} at $n_1 = 800$.
The sample size $n_2$ is $800$ and the noise variance $\sigma^2$ is $1/4$.
%Covariate shift: We set $\kappa = 1$ and $d = 0$.
%We set $\rho_2 = 4$ and vary $\rho_1$ from $5$ to $25$ for sample sizes.
%We use the scale parameter $\lambda = 1$ for the curve without covariate shift and $\lambda = 2$ for the curve with covariate shift (cf. Section \ref{sec_covshift}).

>>>>

\begin{figure*}[!t]
	\centering
	\includegraphics[width=0.3\textwidth,valign=t]{figures/mtl_model_arch.pdf}
	\caption{A hard parameter sharing architecture}
	\label{fig_intro_arch}
\end{figure*}
%\begin{enumerate}
%	\item $X = Z \Sigma^{1/2} \in \real^{n\times p}$ for a positive semidefinite matrix $\Sigma\in \R^{p\times p}$, and every entry of $Z \in \real^{n \times p}$ is drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$.
%	\item every entry of $\varepsilon^{(i)} \in \real^{n \times t}$ is drawn indepdently from a one dimensional distribution with zero mean, variance $\sigma^2$, and bounded moment up to any order.\footnote{There exists a fixed function $C: \mathbb{N} \rightarrow \real^+$ such that for any $k \in \N$, the $k$-th moment is bounded by $C(k)$.}
%\end{enumerate}

@@

\paragraph{Examples.}
Obviously, the concentration error decreases with $n$.
First, we consider the variance of $\hat{\beta}_i^{\MTL}$, which is $\sigma^2\norm{a_i^{\star}} \bigtr{\Sigma (X^{\top} X)^{-1}}$?
It turns out that this quantity converges to a fixed limit in the high-dimensional setting, which is formally stated in the following assumption.

@@

\paragraph{Simulations.}
First, we validate the result of Example \ref{ex_same_cov}.
Figure \ref{fig_same_cov} shows the average prediction loss over ten tasks as we increase the number of samples per-task from $400$ to $2000$.
In all the parameter settings, our results estimate the empirical losses accurately.
We also observe a trend that the average prediction loss increases as we increase distance $d$ from $0.1$ to $0.2$.
Our work explains the differences between these two settings since $d^2 = 0.1^2$ is always smaller than $\frac{\sigma^2 p}{n - p}$, but $d^2 = 0.2^2$ is not.
%Second, when $d = 0.1$, we have that $d^2 \le \frac{\sigma^2 p}{n - p}$ for all values of $n$, hence the average prediction loss of hard parameter sharing is always lower than STL.
Indeed, we observe a crossover point between hard parameter sharing and STL.
Finally, for $d = 0.2$, looking horizontally, we find that HPS requires fewer samples per-task than STL to achieve the same loss level. %\FY{I do not quite understand this sentence about "3x fewer", because how much data needed depends on $d$ and the prediction loss level we are looking at. For example, at the cross point, this ratio is 1. }

%Our result from Corollary \ref{cor_MTL_loss} explains this trend.

%Covariate shift: We set $\kappa = 1$ and $d = 0$.
%We set $\rho_2 = 4$ and vary $\rho_1$ from $5$ to $25$ for sample sizes.
%We use the scale parameter $\lambda = 1$ for the curve without covariate shift and $\lambda = 2$ for the curve with covariate shift (cf. Section \ref{sec_covshift}).

### 2021-06-12 revision_journal/results.tex
restructure

@@

@@

To describe quantitively the bias-variance tradeoff phenomenon and the effect of varying sample sizes on the predication loss, we consider a \emph{random-effect model}, which has been studied for  single-task linear regression and ridge regression (see e.g. \cite{dobriban2020wonder,dobriban2018high}).

@@

>>>>

\paragraph{Examples.}

@@

\paragraph{Simulations.}
Finally, we validate the result of Example \ref{ex_covshift}.
Figure \ref{fig_covariate} shows task two's prediction loss as we increase task one's sample size.
Recall that $\lambda$ measures the severity of covariate shifts---a larger $\lambda$ means a larger covariate shift.
We indeed observe the dichotomy in Example \ref{ex_covshift} at $n_1 = 800$.
The sample size $n_2$ is $800$ and the noise variance $\sigma^2$ is $1/4$.

@@

\paragraph{Examples.}
To describe quantitively the bias-variance tradeoff phenomenon and the effect of varying sample sizes on the predication loss, we consider a \emph{random-effect model}, which has been studied for  single-task linear regression and ridge regression (see e.g. \cite{dobriban2020wonder,dobriban2018high}).

@@

\paragraph{Simulations.}
Second, we validate the result of Example \ref{ex_sample_ratio}.
Figure \ref{fig_size} shows task two's prediction loss  as we increase the sample ratio $n_1 / n_2$ from $1/10$ to $7/10$.
%Again, our estimates are accurate compared to the empirical losses.
We consider a regime where task two consists of $80,000$ samples, and task one's sample size varies from $8,000$ to $56,000$.
The task-specific variance (which scales with model distance) is $d = 0.2$, the noise variance is $\sigma^2 = 4^2$, and the shared signal variance is $1$. We observe that as we increase the sample ratio, task two's prediction loss decreases initially but later will increase when the sample ratio is above a certain level.
On the other hand, when $d = 0.15$, task two's prediction loss decreases faster.
Intuitively, this is because bias increases less for smaller $d^2$.
